A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2021-07-02 Papers/Amlap_2020_talk/amlap2020_abstract.tex
rearrange files

We also investigated the relationship between a word's predictability and its RT. We fit Bayesian regression models using suprisal, frequency, length, and surprisal:length and frequency:length interactions at both the current and previous word as predictors of RT (see Table~1 for details). We use 3 models for surprisals: a Kneser-Ney smoothed 5-gram model, a recurrent neural network model (GRNN) from Gulordava et al. (2018), and a transformer model (TXL) from Dai et al. (2019). Across all models, we see large effects of current word surprisal and length, which noticeably exceed effects of previous word predictors (Table~1). We see minimal to no frequency effects (consistent with Shain 2019). In a nested model comparison, we found GRNN and TXL encoded complementary information, but adding 5-gram surprisals did not improve a regression that already had GRNN. We also fit GAMs to look at the shape of the relationship between surprisal and RT; we find a linear relationship between RT and current word surprisal and no relationship with previous word surprisal (Figure~3). The linear relationship matches known effects from eye-tracking and self-paced reading (Smith \& Levy 2013), but the lack of spillover indicates a potential advantage of Maze over eye-tracking and self-paced reading.

In summary, we provide an adaptation of the Maze task for long naturalistic texts, and show that participants can comprehend what they read during A-maze. Additionally, we find robust, localized surprisal effects, supporting Forster et al.'s (2009) argument that Maze forces highly incremental processing. Overall, this suggests that A-maze is a versatile alternative to eye-tracking and self-paced reading. Code and data at github.com/vboyce/amaze-natural-stories.

>>>>

We also investigated the relationship between a word's predictability and its RT. We fit Bayesian regression models using suprisal, frequency, length, and surprisal:length and frequency:length interactions at both the current and previous word as predictors of RT (see Table~1 for details). We use 3 models for surprisals: a Kneser-Ney smoothed 5-gram model, a recurrent neural network model (GRNN) from Gulordava et al. (2018), and a transformer model (TXL) from Dai et al. (2019). Across all models, we see large effects of current word surprisal and length, which noticeably exceed effects of previous word predictors (Table~1). We see minimal to no frequency effects (consistent with Shain 2019). In a nested model comparison, we found the GRNN fit human reading times the best, although TXL seems to encode some complementary information. We also fit GAMs to look at the shape of the relationship between surprisal and RT; we find a linear relationship between RT and current word surprisal and no relationship with previous word surprisal (Figure~3). The linear relationship matches known effects from eye-tracking and self-paced reading (Smith \& Levy 2013), but the lack of spillover indicates a potential advantage of Maze over eye-tracking and self-paced reading.

In summary, we provide an adaptation of the Maze task for long naturalistic texts, and show that participants can comprehend what they read during A-maze. Additionally, we find robust, localized surprisal effects, supporting Forster et al.'s (2009) argument that Maze forces highly incremental processing. Overall, this suggests that A-maze is a versatile alternative to eye-tracking and self-paced reading. Code and data at \url{github.com/vboyce/amaze-natural-stories}.
